### 2015-06-06 chapters/pitfalls.tex
adding pitfall absence of equivalence

>>>>

% ==========================
\section{Pitfall: Absence of canonical equivalence}
\label{pitfall-absence-of-equivalence}
% ==========================

Although in most cases canonical equivalence will take care of alternative
encodings of the same character, there are a some cases in which the Unicode
Standard decided against equivalence. This leads to identical characters that
are not equivalent, like <ø> \textsc{latin small letter o with stroke} at
\uni{00F8} and <o̷> a combination of \textsc{latin small letter o} at \uni{006F}
with \textsc{combining short solidus overlay} at \uni{0037}.

The general rule followed is that extensions of Latin characters that are
connected to the character are not separated as combining diacritics. For
example, characters like <ƞ ɲ ɳ> or <ɰ ɱ> are obviously derived from <n> and <m>
respectively, but they are treated like new separate characters in the Unicode
Standard. Likewise, characters like <ø> and <ƈ> are not separated into a base 
character <o> and <c> with an attached combining diacritic.

Interestingly, and somewhat illogically, there are three elements attached to
other characters that are still treated as separable in the Unicode Standard. First,
the cedilla, like in <ç> \textsc{latin small letter c with cedilla} at
\uni{00E7}, second the ogonek, like in <ą> \textsc{latin small letter a with
ogonek} at \uni{0105} and, third, the horn, like in <ơ> \textsc{latin small
letter o with horn} at \uni{01A1}. These characters are indeed decomposed in a 
base character with a combining diacritic. However, it is these cases that should be 
considered the exceptions to the rule.

There are two combinations that deserve specific care because it is actually 
possible to produce identical characters in different ways, without them being 
canonical equivalent. First, there are the
combining characters designated as ``overlay'' in the Unicode Standard, like    
<\ \ {\large  ̴}> \textsc{combining tilde overlay} at \uni{0334} or 
<\ \ {\large  ̵}> \textsc{combining short stroke overlay} at \uni{0335}. 
There are many characters that look like they are 
precomposed with such an overlay, for example 
<{\small \fontspec{CharisSIL}ɫ ᵬ ᵭ ᵱ}> or 
<{\small \fontspec{CharisSIL}ƚ ɨ ɉ ɍ}>, or also the 
example of <ø> given at the start of this section. Second, the same situation also 
occurs with combining characters designated as ``hook'', like < {\large  ̡}> 
\textsc{combining palatalized hook below} at \uni{0321}. This element seems to
occur in precomposed characters like <{\small \fontspec{CharisSIL}ᶀ ᶁ ᶂ ᶄ}>. 
However, all these 
characters are not canonical equivalent with a combination of their base 
characters with the combining overlay or combining hook.
